---
title: "Phishing with SQL"
description: |
  Welcome to the website. I hope you enjoy it!
site: distill::distill_website
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)
library(dplyr)
library(rmarkdown)
library(plotly)
library(ggplot2)
library(distill)

```


We started with [fake news](https://zhusophia.github.io/Fake_Fake_News/), only to end up at phishing. That's just how the world works! 

Taking raw data from [this dataset](https://www.kaggle.com/datasets/danielfernandon/web-page-phishing-dataset?resource=download), I'm attempting to explore the URLs of phishing websites! 

Taking a look at the raw data: 

DISPLAY RAW DATA 
- grab a first 4 columns + phishing column (move phsidhing column tot he front)


```{r load data}
data <- read.csv("web-page-phishing.csv")
```

```{r general data}


data <- data %>% 
  select(phishing, everything())

paged_table(head(data))

```
<!-- ggplot(phishingpie, aes(x="", y=occurances, fill=phishing)) +
  geom_col() +
  coord_polar(theta="y")+
  theme_void()


-->

```{r phishing pie chart}

phishingpie <- data %>%
  select(phishing) %>%
  group_by(phishing) %>% 
  summarise(occurances = n()) %>%
  mutate(phishing = recode(phishing, '1'='Phishing link', '0'="Non-phishing link"))

phishingagg <- plot_ly(phishingpie, labels = ~phishing, values = ~occurances, type = 'pie',marker = list(colors = c('MidnightBlue', 'MediumPurple')),
        textposition = 'inside',
        textinfo = 'label+percent',
        insidetextfont = list(color = '#FFFFFF'),
        hoverinfo = 'text',
        text = ~paste(occurances, 'links')) %>% 
  layout(title="Phishing V.S. Non-Phishing Links", 
         legend = list(x = 100, y = 0.5, title=list(text='<b> Types of Websites </b>')))

phishingagg

```


<!--




ggplot(urldata, aes(x=url_length, color=phishing)) + 
  geom_density()
---> 
```{r url_length m_and_sd}

pmean <- data %>% 
  select( phishing, url_length) %>%
  filter(phishing==1) %>%
  summarise(mean_p = mean(url_length))

psd <- data %>% 
  select( phishing, url_length) %>%
  filter(phishing==1) %>%
  summarise(sd_p = sd(url_length))

npmean <- data %>% 
  select( phishing, url_length) %>%
  filter(phishing==0) %>%
  summarise(mean_np = mean(url_length))

npsd <- data %>% 
  select( phishing, url_length) %>%
  filter(phishing==0) %>%
  summarise(sd_np = sd(url_length))

data.frame(pmean, psd, npmean, npsd)


```
```{r urldensity}

urldensity <- data.frame(
  phishingurl=factor(rep(c("P", "NP"), each=300)),
  weight=round(c(rnorm(300, mean=66.49112, sd=68.59874),
                 rnorm(300, mean=23.58997	, sd=16.13607)))
  )

df <- data %>% 
  select( phishing, url_length)


p <- ggplot(urldensity, aes(x=weight, color=phishingurl)) +
  geom_density() +
  geom_vline(aes(xintercept=66.49112),
             linetype="dashed", color='blue') +
  geom_vline(aes(xintercept=23.58997),
             linetype="dashed", color='blue') 
  p + scale_x_continuous(expand = c(0, 0), limits = c(0, NA)) + 
  scale_y_continuous(expand = c(0, 0), limits = c(0, NA)) 

```
- make the links open to a blank page 

```{sql test, eval=FALSE, echo=T} 
can you display just this as something
```


- general talk about teh data 
-> how many were fishing? how many weren't? -> pie chart 

correlation 
- build tables (take the head of x)
- explain the pearson correlation, how can we get info out 
- cast bigint and basically couldn twork, so had to build a table (you can build it using R)
create table 


url length 
- out of phishing, plot density graph, 
--> url lenght vs amount of urls -> phishing vs no phishing 

same thing with redirections 

- end

this shoudl be pretty quick and simple it hink

https://docs.google.com/spreadsheets/d/1JNzGqUo9HEn5WOKftEgcm94JddOK6G3_QW-UI4sFNUY/edit#gid=0

https://docs.google.com/document/d/1M9kvGEJ1_db81M5g6J8DtF0kAFnCCAPlnuUMzAmVUOA/edit
