---
title: "Phishing with SQL"
description: |
  Welcome to the website. I hope you enjoy it!
site: distill::distill_website
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)
library(dplyr)
library(plotly)
library(ggplot2)
library(distill)

```


We started with [fake news](https://zhusophia.github.io/Fake_Fake_News/), only to end up at phishing. That's just how the world works! 

Taking raw data from [this dataset](https://www.kaggle.com/datasets/danielfernandon/web-page-phishing-dataset?resource=download), I'm attempting to explore the URLs of phishing websites! 

Taking a look at the raw data: 

DISPLAY RAW DATA 
- grab a first 4 columns + phishing column (move phsidhing column tot he front)


```{r, echo=TRUE}
library(rmarkdown)
paged_table(mtcars)
```


- make the links open to a blank page 

```{sql test, eval=FALSE, echo=T} 
can you display just this as something
```


- general talk about teh data 
-> how many were fishing? how many weren't? -> pie chart 

correlation 
- build tables (take the head of x)
- explain the pearson correlation, how can we get info out 
- cast bigint and basically couldn twork, so had to build a table (you can build it using R)
create table 


url length 
- out of phishing, plot density graph, 
--> url lenght vs amount of urls -> phishing vs no phishing 

same thing with redirections 

- end

this shoudl be pretty quick and simple it hink

https://docs.google.com/spreadsheets/d/1JNzGqUo9HEn5WOKftEgcm94JddOK6G3_QW-UI4sFNUY/edit#gid=0

https://docs.google.com/document/d/1M9kvGEJ1_db81M5g6J8DtF0kAFnCCAPlnuUMzAmVUOA/edit
