---
title: "Phishing with SQL"
description: |
  Welcome to the website. I hope you enjoy it!
site: distill::distill_website
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)
library(dplyr)
library(rmarkdown)
library(plotly)
library(ggplot2)
library(distill)

```


We started with [fake news](https://zhusophia.github.io/Fake_Fake_News/), only to end up at phishing. That's just how the world works! 

Taking raw data from [this dataset](https://www.kaggle.com/datasets/mikhail1681/walmart-sales?resource=download), I'm looking at sales trends for the year  

Spanning just under two yeras, this dataset keeps tracks of the weekly sales of certain Walmart stores. there are 45 stores and 10 holidays. it also takes general indicators of the wellbeing of economy and enviormnet - temprarture, CPI, unemployement, and fuel prices.

```{python import, eval=FALSE, echo=T}
import pandas as pd
df = pd.read_csv('Walmart_sales.csv', parse_dates=['Date'], dayfirst=False)

```

```{python cleaning, eval=FALSE, echo=T}
#sort by store number and then date
df.sort_values(by=['Store', 'Date'])

#standarize dates to mm-dd-yyyy
df['Date'] = pd.to_datetime(df['Date'], format='%d-%m-%Y').dt.strftime('%m-%d-%Y')

#rounding data 
df.round({'Weekly_Sales':2, 'Temperature':0, 'Fuel_Price':2, 'CPI':3, 'Unemployment':3})

#removing missing values
df.dropna

#create cleaned csv
df.to_csv('./WalmartSalesClean.csv')

```


after cleaning, the information appears as so: 


```{r load data2}
data <- read.csv("WalmartSalesClean.csv")
```

```{r general1 data}
paged_table(head(data))

```

Let's answer a few questions about the data. We'll be using SQL to clean and manipulate the data. 


1. Which holidays affect weekly sales the most?


```{sql test1, eval=FALSE, echo=T}
ALTER TABLE WalmartSalesClean
ALTER COLUMN Holiday_Flag Int;

SELECT Date, SUM(CAST(Weekly_Sales AS FLOAT)) AS Weekly_Sales
FROM WalmartSalesClean 
WHERE Holiday_Flag='1'
GROUP BY Date
ORDER BY Weekly_Sales DESC;

SELECT Holiday_Flag = '1', Date, SUM(CAST(Weekly_Sales AS FLOAT)) AS Weekly_Sales
FROM WalmartSalesClean
GROUP BY Date
ORDER BY Weekly_Sales DESC;
```

2. Which stores in the dataset have the lowest and highest unemployment rate?  What factors do you think are impacting the unemployment rate?
```{sql test2, eval=FALSE, echo=T}
WITH corrdata AS (
SELECT Store, avg(CAST(Unemployment AS decimal)) AS Unemployment, avg(CAST(Temperature AS decimal)) AS Temperature, avg(CAST(Fuel_Price AS decimal)) AS Fuel_Prices, avg(CAST(CPI AS decimal)) AS CPI, avg(CAST(Weekly_Sales AS decimal)) AS Weekly_Sales
FROM WalmartSalesClean
GROUP BY Store
/*ORDER BY Unemployment desc, Weekly_Sales */

```

Pearson's correlation
```{sql test3, eval=FALSE, echo=T}
rawdata AS (
SELECT Unemployment AS x, Weekly_Sales AS y, Unemployment*Weekly_Sales AS xy, Unemployment*Unemployment AS xsquared, Weekly_Sales*Weekly_Sales AS ysquared	
FROM corrdata
),

pdata AS (
SELECT SUM(x) AS sumx, SUM(y) AS sumy, SUM(xy) as sumxy, SUM(xsquared) as sumx2, SUM(ysquared) as sumy2, COUNT(x) as n
FROM rawdata
)

SELECT (n*sumxy - sumx * sumy) / (sqrt((n*sumx2 - sumx*sumx)*(n*sumy2 - sumy*sumy))) AS "Pearson's Correlation"
FROM pdata

```

3. Is there any correlation between CPI and Weekly Sales?  How does the correlation differ when the Holiday Flag is 0 versus when the Holiday Flag is 1?
```{sql test4, eval=FALSE, echo=T}
WITH rdata AS (
SELECT CAST(ROUND(CPI, 0) AS INT) AS CPI, CAST(Weekly_Sales AS decimal) AS Weekly_Sales
FROM WalmartSalesClean
WHERE Holiday_Flag = 0
),

corrdata AS(
SELECT CPI, AVG(Weekly_Sales) AS Weekly_Sales
FROM rdata
GROUP BY CPI
/*ORDER BY CPI ASC*/
),

```

```{sql test5, eval=FALSE, echo=T}
rawdata AS (
SELECT CPI AS x, Weekly_Sales AS y, CPI*Weekly_Sales AS xy, CPI*CPI AS xsquared, Weekly_Sales*Weekly_Sales AS ysquared	
FROM corrdata
),

pdata AS (
SELECT SUM(x) AS sumx, SUM(y) AS sumy, SUM(xy) as sumxy, SUM(xsquared) as sumx2, SUM(ysquared) as sumy2, COUNT(x) as n
FROM rawdata
)

SELECT (n*sumxy - sumx * sumy) / (sqrt((n*sumx2 - sumx*sumx)*(n*sumy2 - sumy*sumy))) AS "Pearson's Correlation"
FROM pdata

```



https://docs.google.com/spreadsheets/d/1JNzGqUo9HEn5WOKftEgcm94JddOK6G3_QW-UI4sFNUY/edit#gid=0